---
title: "Introduction to R Statistical Analysis Portfolio"
author: "Gabriel Demetrios Lafis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to R Statistical Analysis Portfolio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center"
)
```

## Overview

The `rstatportfolio` package provides a comprehensive framework for statistical analysis in R. It includes tools for exploratory data analysis, statistical modeling, time series forecasting, and interactive visualization. This vignette provides an introduction to the package and demonstrates its key features.

## Installation

You can install the development version of `rstatportfolio` from GitHub:

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("galafis/r-statistical-analysis-portfolio")
```

## Loading the Package

```{r setup}
library(rstatportfolio)
```

## Sample Data

The package includes sample datasets for demonstration purposes:

- `regression_data`: A dataset for regression and classification tasks
- `time_series_data`: A dataset for time series analysis
- `time_series_ts`: A time series object

Let's load the sample data:

```{r eval=FALSE}
data(regression_data)
data(time_series_data)
data(time_series_ts)
```

For this vignette, we'll use the built-in `mtcars` dataset:

```{r}
# View the first few rows of mtcars
head(mtcars)
```

## Exploratory Data Analysis

The `perform_eda()` function provides a comprehensive exploratory data analysis:

```{r eval=FALSE}
# Perform EDA on mtcars
eda_results <- perform_eda(
  data = mtcars,
  target = "mpg",
  plots = TRUE
)

# View summary of EDA results
summary(eda_results)
```

The EDA results include:

- Data summary statistics
- Variable distributions
- Correlation analysis
- Target variable relationships
- Visualizations

## Statistical Modeling

The package provides functions for building and comparing multiple statistical models:

```{r eval=FALSE}
# Build regression models
model_results <- build_regression_models(
  data = mtcars,
  target = "mpg",
  predictors = c("cyl", "disp", "hp", "wt", "qsec"),
  model_types = c("linear", "random_forest"),
  test_size = 0.2,
  seed = 123
)

# View model comparison
summary(model_results)

# Plot model comparison
plot(model_results)
```

For classification tasks:

```{r eval=FALSE}
# Create a binary target
mtcars$am_factor <- factor(mtcars$am, labels = c("automatic", "manual"))

# Build classification models
class_results <- build_classification_models(
  data = mtcars,
  target = "am_factor",
  predictors = c("mpg", "cyl", "hp", "wt"),
  model_types = c("logistic", "random_forest"),
  test_size = 0.2,
  seed = 123
)

# View model comparison
summary(class_results)
```

## Time Series Analysis

For time series analysis, the package provides the `analyze_time_series()` function:

```{r eval=FALSE}
# Load AirPassengers dataset
data(AirPassengers)

# Analyze time series
ts_results <- analyze_time_series(
  data = AirPassengers,
  h = 24,  # Forecast 24 periods ahead
  models = c("auto", "ets"),
  seasonal = TRUE
)

# View forecast
plot(ts_results, "forecast")

# View decomposition
plot(ts_results, "decomposition")

# View ACF and PACF
plot(ts_results, "acf")
plot(ts_results, "pacf")
```

## Conclusion

The `rstatportfolio` package provides a unified framework for statistical analysis in R. It simplifies the process of exploratory data analysis, statistical modeling, and time series forecasting. The package is designed to be user-friendly while providing powerful analytical capabilities.

For more detailed examples and advanced usage, please refer to the other vignettes:

- `vignette("exploratory-data-analysis")`
- `vignette("statistical-modeling")`
- `vignette("time-series-analysis")`

## References

- R Core Team (2023). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
- Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). "Welcome to the tidyverse." Journal of Open Source Software, 4(43), 1686. doi:10.21105/joss.01686.
- Hyndman RJ, Khandakar Y (2008). "Automatic time series forecasting: the forecast package for R." Journal of Statistical Software, 26(3), 1-22. URL http://www.jstatsoft.org/article/view/v027i03.

